Electro-Drop Bouncing in Low-Gravity

### 2018-10-02 doc/paper/aiptemplate.tex
large revision to paper

This electro-drop bounce phenomenon can be extended to useful applications in a space environment, generally to tackle various manifestations of the so-called ``phase separation'' problem of separating a gas phase from a multi-phase flow (or the reverse case), in low-gravity. Removal of satellite droplets produced during pipetting in wet-lab research outside of a glovebox environment aboard ISS has also been recently suggested as a application of the work [private communication, NASA JSC, E. Unger, 2017]. Drops can become spontaneously charged by contact with standard micropipette tips \citep{choi_spontaneous_2013}, and this free charge can possibly be leveraged for the purposes of phase separation in low-gravity.

@@

\subsection{Model Validation}
Using the parameter estimates we find $\mathbb{E}\mbox{u}$ for each drop jump. Dimensional drop apoapses shown in Figure \ref{fig:series_s_eu} scale with $\mathbb{E}\mbox{u}$ as expected according to our earlier analysis. Electrostatic Euler numbers in the data set vary between $1.4 \lesssim \mathbb{E}\mbox{u} \lesssim 35.4$. \sout{The dielectrophoretic force plays a small role when drops have net charge in a DC field. The condition to neglect the DEP force was satisfied for all experiments in the dataset with typical values of the condition number $\kappa_w \epsilon_0 K R_d^2 E_0/q \approx \mathcal{O}(10^{-6})$}. In the non-dimensional trajectories with short-time scaling shown in Figure \ref{fig:series_s_ds}, we see that the scaled trajectory apoapses are consistently $\mathcal{O}(1)$, with all trajectories overshooting their characteristic time scale (which predicts returns at $t^*  =2$ at zeroth order). A trend of increasing $\mbox{max} \left(y^* \right)$ at apogee for decreasing $\mathbb{E}\mbox{u}$ is seen in Figure \ref{fig:yscale_trend}. This trend seems to decay to a constant $y_{max}/y_c \approx 1/2$ for $6 < \mathbb{E}\mbox{u} < 10$, and implies the existence an intermediate regime with a larger characteristic lengthscale for $\mathbb{E}\mbox{u} < 6$. For $\mathbb{E}\mbox{u} > 10$ the trend reverses and apoapses begin to grow again.
\begin{figure}[!htb]
    \centering
    \resizebox{0.5\textwidth}{!}{\input{../figures/series.pgf}}
    \caption{Drop trajectories as a function of $\mathbb{E}\mbox{u}$.\label{fig:series_s_eu}}
\end{figure}
\begin{figure}[htb]
    \centering
    \resizebox{0.5\textwidth}{!}{\input{../figures/series_s_ds.pdf_tex}}
    \caption{Non-dimensional trajectories with the short-time scaling.\label{fig:series_s_ds}}
\end{figure}
\begin{figure}[htb]
    \centering
    \resizebox{0.5\textwidth}{!}{\input{../figures/yscale_trend.pgf}}
    \caption{Non-dimensional apoapse height $y_{max}/y_c$ compared with $\mathbb{E}\mbox{u}$.\label{fig:yscale_trend}}
\end{figure}

Because $\mathbb{E}\mbox{u} \ll 1$ is not satisfied for any of the experimental drops the entire dataset is beyond the regime where the asymptotic results for short-times, given by Equation \ref{perturb_image}, are valid. However, the long-time scaled asymptotic trajectory  and time-of-flight $t_f$ given by Equations \ref{perturb_viscous} and \ref{time_of_flight}, when redimensionalized by the short-time characteristic time $t_c=\mathbb{E}\mbox{u} R_d$, compare favorably to the experimental trajectory apoapses and time-of-flight $t_b$ as shown in Figures \ref{fig:times2} and \ref{fig:ymaxes}. This allows an improvement to be made to the asymptotic time-of-flight result of Equation \ref{time_of_flight} by multiplying the series by the empirical coefficient $a = 1.31$. The semi-analytic time-of-flight is then given by
\begin{equation}
\label{time_improved}
t^* = 2.62 + 1.75\phi\mathbb{E}\mbox{u} + 1.05\phi^2\mathbb{E}\mbox{u}^{2} + \mathcal{O}(\phi^3\mathbb{E}\mbox{u}^{3}).
\end{equation}
\begin{figure}[htb]
    \centering
    \resizebox{0.5\textwidth}{!}{\input{../figures/times.pgf}}
    \caption{Dimensional time-of-flight asymptotic estimates $t_c t_f$ compared with experimental time-of-flight $t_b$. These results compare $\mathcal{O}(1) \equiv y_c$ and $\mathcal{O}(\phi^3\mathbb{E}\mbox{u}^{3})$ accurate variations of Equation \ref{time_of_flight}. \label{fig:times2}}
\end{figure}
\begin{figure}[htb]
    \centering
    \resizebox{0.5\textwidth}{!}{\input{../figures/ymaxes.pgf}}
    \caption{Dimensional asymptotic trajectory apoapses, found from Equation \ref{perturb_viscous} with $t^*=1$, as compared to experimental apoapses $y_{max}$.\label{fig:ymaxes}}
\end{figure}
A similar empirical coefficient $a = 1.31$ is found for Equation \ref{perturb_viscous}. 
The relative magnitudes of the forces acting on the drops determined by the MLE numerical solutions to Equation \ref{gov_eqn_subs} for the entire population of drop tower experiments are shown in Figure \ref{fig:forces}. Coulomb, image, and drag forces acting on the drops vary in typical magnitude between $\mathcal{O}(10^{-6})$-$\mathcal{O}(10^{-4})$ N. We see that of the drops in the experimental dataset only the two with the largest $\mathbb{E}\mbox{u} \sim \mathcal{O}(10)$ could appropriately be said to be in the inertial electro-viscous regime. In all other cases image forces are much stronger than drag. As expected, the image forces themselves rapidly become small compared to Coulomb forces for drops with apoapses $\mbox{max}\left( y\right) \gtrsim L$. The drop with largest $\mathbb{E}\mbox{u}$ in our dataset failed to escape the electric field as the escape condition $\phi \mathbb{E}\mbox{u} / 8\pi = 0.2 < 1$ was unsatisfied. Equation \ref{time_improved} predicts that this drop will return to the substrate after 7.4 s, a period of free-fall which is lamentably well out of reach of the Dryden Drop Tower. However, such an experiment could be performed in a drop tower facility with a longer free fall period, aboard the ISS or on certain suborbital parabolic flights.
\begin{figure}[!htb]
    \centering
    \resizebox{0.5\textwidth}{!}{\input{../figures/forces_all.pgf}}
    \caption{Simulated forces acting on drops up to the first apoapse as determined by the MLE solutions to Equation \ref{gov_eqn_subs} for all drops in the dataset. The (\protect\redline) lines denote the Coulomb force, the (\protect\blueline) lines the drag force, and the (\protect\cyanline) lines the foce due to image charge respectively. The forces are scaled by the drop inertia.\label{fig:forces}}
\end{figure}

>>>>

This electro-drop bounce phenomenon may have useful applications in a space environment, generally to tackle various manifestations of the so-called ``phase separation'' problem of separating a gas phase from a multi-phase flow (or the reverse case), in low-gravity. Removal of satellite droplets produced during pipetting in wet-lab research outside of a glovebox environment aboard ISS has been recently suggested as a application of the work [private communication, NASA JSC, E. Unger, 2017]. Drops can become spontaneously charged by contact with standard micropipette tips \citep{choi_spontaneous_2013}, and this free charge can possibly be leveraged for the purposes of phase separation in low-gravity.

@@

\subsection{Model Validation}
Using the parameter estimates we find $\mathbb{E}\mbox{u}$ for each drop jump. Dimensional drop apoapses shown in Figure \ref{fig:series_s_eu} scale with $\mathbb{E}\mbox{u}$ as expected according to our earlier analysis. Electrostatic Euler numbers in the data set vary between $1.4 \lesssim \mathbb{E}\mbox{u} \lesssim 35.4$. \sout{The dielectrophoretic force plays a small role when drops have net charge in a DC field. The condition to neglect the DEP force was satisfied for all experiments in the dataset with typical values of the condition number $\kappa_w \epsilon_0 K R_d^2 E_0/q \approx \mathcal{O}(10^{-6})$}. In the non-dimensional trajectories with short-time scaling shown in Figure \ref{fig:series_s_ds}, we see that the scaled trajectory apoapses are consistently $\mathcal{O}(1)$, with all trajectories overshooting their characteristic time scale (which predicts returns at $t^*  =2$ at zeroth order). A trend of increasing $\mbox{max} \left(y^* \right)$ at apogee for decreasing $\mathbb{E}\mbox{u}$ is seen in Figure \ref{fig:yscale_trend}. This trend seems to decay to a constant $y_{max}/y_c \approx 1/2$ for $6 \lesssim \mathbb{E}\mbox{u} \lesssim 20$, and implies the existence an intermediate regime with a larger characteristic lengthscale for $\mathbb{E}\mbox{u} \lesssim 6$.
\begin{figure}[!htb]
    \centering
    \resizebox{0.5\textwidth}{!}{\input{../figures/series.pgf}}
    \caption{Drop trajectories as a function of $\mathbb{E}\mbox{u}$.\label{fig:series_s_eu}}
\end{figure}
\begin{figure}[htb]
    \centering
    \resizebox{0.5\textwidth}{!}{\input{../figures/series_s_ds.pdf_tex}}
    \caption{Non-dimensional trajectories with the short-time scaling.\label{fig:series_s_ds}}
\end{figure}
\begin{figure}[htb]
    \centering
    \resizebox{0.5\textwidth}{!}{\input{../figures/yscale_trend.pgf}}
    \caption{Non-dimensional apoapse height $y_{max}/y_c$ compared with $\mathbb{E}\mbox{u}$.\label{fig:yscale_trend}}
\end{figure}

Because $\mathbb{E}\mbox{u} \ll 1$ is not satisfied for any of the experimental drops the entire dataset is beyond the regime where the asymptotic results for short-times, given by Equation \ref{perturb_image}, are valid. However, the long-time scaled asymptotic trajectory   $y_c$ and time-of-flight $t_f$ given by Equations \ref{perturb_viscous} and \ref{time_of_flight}, when redimensionalized by the short-time characteristic time $t_c=\mathbb{E}\mbox{u} R_d$, compare favorably to the experimental trajectory apoapses $y_{max}$ and time-of-flight $t_b$ as shown in Figures \ref{fig:times2} and \ref{fig:ymaxes}. This allows an improvement to be made to the asymptotic time-of-flight result of Equation \ref{time_of_flight} by multiplying the series by the empirical coefficient $a = 1.31$. The semi-analytic time-of-flight is then given by
\begin{equation}
\label{time_improved}
t^* = 2.62 + 1.75\phi\mathbb{E}\mbox{u} + 1.05\phi^2\mathbb{E}\mbox{u}^{2} + \mathcal{O}(\phi^3\mathbb{E}\mbox{u}^{3}).
\end{equation}
\begin{figure}[htb]
    \centering
    \resizebox{0.5\textwidth}{!}{\input{../figures/times.pgf}}
    \caption{Dimensional time-of-flight asymptotic estimates $t_c t_f$ compared with experimental time-of-flight $t_b$. These results compare $\mathcal{O}(1) \equiv 2 t_c$ and $\mathcal{O}(\phi^3\mathbb{E}\mbox{u}^{3})$ accurate variations of Equation \ref{time_of_flight}. \label{fig:times2}}
\end{figure}
\begin{figure}[htb]
    \centering
    \resizebox{0.5\textwidth}{!}{\input{../figures/ymaxes.pgf}}
    \caption{Dimensional asymptotic trajectory apoapses, found from Equation \ref{perturb_viscous} with $t^*=1$ and $\mathcal{O}(1) \equiv y_c/2$, as compared to experimental apoapses $y_{max}$. \label{fig:ymaxes}}
\end{figure}
A similar empirical coefficient $a = 1.22$ is found for Equation \ref{perturb_viscous}. 
The relative magnitudes of the forces acting on the drops determined by the MLE numerical solutions to Equation \ref{gov_eqn_subs} for the entire population of drop tower experiments are shown in Figure \ref{fig:forces}. Coulomb, image, and drag forces acting on the drops vary in typical magnitude between $\mathcal{O}(10^{-6})$-$\mathcal{O}(10^{-4})$ N. We see that of the drops in the experimental dataset only the two with the largest $\mathbb{E}\mbox{u} \sim \mathcal{O}(10)$ could appropriately be said to be in the inertial electro-viscous regime. In all other cases image forces are much stronger than drag. As expected, the image forces themselves rapidly become small compared to Coulomb forces for drops with apoapses $\mbox{max}\left( y\right) \gtrsim L$. The drop with largest $\mathbb{E}\mbox{u}$ in our dataset failed to escape the electric field as the escape condition $\phi \mathbb{E}\mbox{u} / 8\pi = 0.2 < 1$ was unsatisfied. Equation \ref{time_improved} predicts that this drop will return to the substrate after 7.4 s, a period of free-fall which is lamentably well out of reach of the Dryden Drop Tower. However, such an experiment could be performed in a drop tower facility with a longer free fall period, aboard the ISS or on certain suborbital parabolic flights.
\begin{figure}[!htb]
    \centering
    \resizebox{0.5\textwidth}{!}{\input{../figures/forces_all.pgf}}
    \caption{Simulated forces acting on drops up to the first apoapse as determined by the MLE solutions to Equation \ref{gov_eqn_subs} for all drops in the dataset. The (\protect\redline) lines denote the Coulomb force, the (\protect\blueline) lines the drag force, and the (\protect\cyanline) lines the foce due to image charge respectively. The forces are scaled by the drop inertia.\label{fig:forces}}
\end{figure}
